Community-Conditioned Language Models

### 2020-11-23 paper/paper.tex
intro tweaks

\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\section{Introduction}

Linguistic communication requires that speakers share 
knowledge of certain linguistic conventions, such as syntactic structure, 
word meanings, and accepted patterns of interaction.
Speakers assume that these conventions are \emph{common ground} among their interlocutors,
based on joint membership in a community \cite{Stalnaker2002, Clark1996}.
Such a community may be of any size,
from the very small, like members of a particular friend group,
to the very large, like speakers of English.
Because it is the source of linguistic convention, the \emph{speech community} 
is an important concept in the study of linguistic variation \citep{Gumperz1972}.

Not only is linguistic variation an object of inquiry in its own right,
but also important to consider in the development of robust, equitable language technology.
Previous work has shown that by taking demographic information into account, 
models can achieve better performance on various standard NLP tasks \citep{Hovy2015,Yang2017},
but little is known about how variation at the level of speech community affects performance.

There is a fair amount of previous work that investigates linguistic variation using computational methods.
However, most of this work considers variation across macro-social categories,
such as gender \citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, 
and geographic location \citep{Eisenstein2010,Bamman2014a}.
An exception to this is \citet{DelTredici2017},
who use the modified skip-gram model of \citet{Bamman2014a} to show that lexical semantic
variation occurs across different communities organized around the same topic.

In the present work, we continue this line of 
investigating linguistic variation across online communities.
In contrast to previous work, we investigate variation across speech communities
(rather than macro-social categories), and use context-sensitive sequence models
capable of capturing variation beyond the lexical level.

Broadly construed, we want to investigate two questions.  First, we
are interested in variation in the level of linguistic complexity
exhibited by different communities.

\jp{
  Some examples (perhaps include if there is space):

@@

  (Videos)
  Reminds me of Y2K computer bug. A lot of work went into
  pre-emptively fixing it, and a lot of bugs were found and fixed,
  some of which would have been catastrophic. Unfortunately, because
  the work was successful, a lot of people now describe it as a waste
  of time/money and a big deal made out of nothing.  }

It is folklore that certain communities exhibit more linguistic complexity than others.
Here we want to quantify this phenomenon.
More precisely, we investigate if the degree of complexity can be
captured in a computational model.  For this purpose define various
Community Conditioned Language Models (CCLM for short), which can be
modeled as an estimator $M$ of the degree to which a given message $m$ 
is acceptable in a given community $c$: \(P_M(m \mid c)\).

Second, we want to find out how, and how much, CCLMs are attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of communities based only on
co-occurrence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language become and the easier it is to identify.

To sum up we aim at testing the following hypotheses:
\begin{hypotheses}
\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
  complexity
\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
  in a CCLM influences its perplexity.
\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
  latent Bayesian classification.
% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
%   influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
  linguistically rich communities.
\end{hypotheses}

>>>>

\begin{document}
\maketitle
\jp{Title suffers from wizwoz system syndrome.
  http://www.cs.sjtu.edu.cn/~kzhu/writing-a-paper-slides.pdf (slide 10+)

  Try to say what the paper brings to the state of the art rather than how it does it.
}
\begin{abstract}
  Using Community-Conditioned Lanaguage Models (CCLMs), we investigate
  the following questions: 1. Can the level of linguistic complexity
  exhibited by different communities is captured. 2. Are CCLMs
  well-tuned to various communities both by intrinsic (can we build a
  community classifier based on CCLMs) and extrinsic methods (are
  CCLMs representations correlated with other network-based
  representations of communities).
\end{abstract}
\jp{Not so good abstract}

\section{Introduction}
\jp{I am seeing computer modern rather than times when building the
  paper!?}  Linguistic communication requires that speakers share
knowledge of certain linguistic conventions, such as syntactic
structure, word meanings, and accepted patterns of interaction.
Speakers assume that these conventions are \emph{common ground} among
their interlocutors, based on joint membership in a community
\cite{Stalnaker2002, Clark1996}.  Such a community may be of any size,
from the very small, like members of a particular friend group, to the
very large, like speakers of English.  Because it is the source of
linguistic convention, the \emph{speech community} is an important
concept in the study of linguistic variation \citep{Gumperz1972}.

Not only is linguistic variation an object of inquiry in its own
right, but it is also important to consider it in the development of
robust, equitable language technology.  Previous work has shown that
by taking demographic information into account, models can achieve
better performance on various standard NLP tasks
\citep{Hovy2015,Yang2017}, but little is known about how variation at
the level of speech community affects performance.
%
There is a fair amount of work that investigates linguistic variation
using computational methods.  However, most of it considers variation
across macro-social categories, such as gender
\citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
geographic location \citep{Eisenstein2010,Bamman2014a}.
\citet{DelTredici2017} provide an exception, by using the modified
skip-gram model of \citet{Bamman2014a} to show that lexical semantic
variation occurs across different communities organized around the
same topic.

In the present work, we further investigate linguistic variation
across online communities. In contrast to previous work, we
investigate variation across speech communities (rather than
macro-social categories), and use context-sensitive sequence models
capable of capturing variation beyond the lexical level.
%
We do so by investigate two questions.  First, we are interested in
variation in the level of linguistic complexity exhibited by different
communities.
\jp{
  Some examples (perhaps include if there is space):

@@

  (Videos)
  Reminds me of Y2K computer bug. A lot of work went into
  pre-emptively fixing it, and a lot of bugs were found and fixed,
  some of which would have been catastrophic. Unfortunately, because
  the work was successful, a lot of people now describe it as a waste
  of time/money and a big deal made out of nothing.  }
%
Indeed, it is folklore that certain communities exhibit more
linguistic complexity than others.  Here we want to quantify this
phenomenon.  More precisely, we investigate if the degree of
complexity can be captured in a computational model.  For this purpose
define various Community Conditioned Language Models (CCLM for short),
which can be modeled as an estimator $M$ of the degree to which a
given message $m$ is acceptable in a given community $c$:
\(P_M(m \mid c)\).

Second, we want to find out how, and how much, CCLMs are attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of communities based only on
co-occurrence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language becomes and the easier it is to identify.

To sum up, we test the following hypotheses:
\begin{hypotheses}
\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
  complexity
\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
  in a CCLM influences its perplexity.
\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
  latent Bayesian classification.
% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
%   influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
  linguistically rich communities.
\end{hypotheses}
